Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields

### 2020-06-04 main_new.tex
Updates from Overleaf

Note that this strategy gives a sequential conditional version of
expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
the expectation is conditional to this data. A similar closed-form calculation
for EIBV is hence applicable in Eq. (\ref{critSEQ}) using the
updated GP model from stage $s-1$. Once the data is collected for the
best design at stage $s$, the GP model is updated again. The mean $\bbm^{s}$ and
covariance matrix $\Sigma^{s}$ are used to compute the next design at
stage $s+1$, and so on.

@@

We now extend the myopic strategy to a look-ahead strategy, which is
optimal when one can gather only two more stages of measurements. In
addition to the next stage of measurements $\by^{j,s}$, this
look-ahead strategy, anticipates the subsequent design indexed $j_2$ with
data $\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$.  The
selected design is:
\begin{eqnarray}\label{critLA}
    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left[ \mbox{argmin}_{j_2}  V_{j_2} \right] \right\}, \\
V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1})\left( 1-p(\x;\mathcal{Y}^{j,s+1})\right) \right) d\mes(u), \nonumber \\
    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
\end{eqnarray}
Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$ represent
the sets of data variables, and the expectations are calculated with respect to
the conditional densities given data at that stage, meaning we have two levels of intermixed minimizations and integrals.
We solve the
first expectation by Monte Carlo sampling of data $\by_{d^{j,s}}$ from its
conditional distribution. For each of these data samples, the second expectation
is solved using the closed-form expressions for EIBV from
Section \ref{sec:ESEP}.

@@

For a fixed survey length, a closed-form expression for the EIBV is available as in Eq. \eqref{eq:eibv}. However, for the sequential approaches this is not the
case. For comparison the properties are therefore evaluated using
Monte Carlo sampling over several replicates of realizations from the model while conducting simulated sequential surveys for each
one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
gathered along the north-south survey lines in these figures. In the
Monte Carlo replicates, such results are repeatedly computed to
approximate the EIBV. We also compare predictive performance
measured by root mean square error (RMSE) for temperature and salinity
estimates as well as also the variance reduction in these two variables. It is
important to note that the objective function used by the agent is
focused on reducing the EIBV, but we nevertheless expect that
we will achieve good predictive performance for criteria such as RMSE
as well. Another non-statistical criteria that is important for
practical purposes is the computational time needed for the strategy,
as this will impact the performance for an embedded system.

@@

We studied the sensitivity of the results by modifying the input parameters to have different correlations between temperature and salinity, standard deviations, and spatial correlation range.  In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of IBV, and much better than \textit{naive}. The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small initial standard deviation or very large spatial correlation range. The \textit{static\_north} design continues to be the best static design for IBV, while \textit{static\_zigzag} is the best design for the other predictive performance measures, especially so with large spatial correlation range. We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the IBV results are not much worse when only temperature data are available. In addition to the comparison made in Table \ref{tab:sim_rhoab}, the current setting includes spatial correlation and this likely reduces the additional influence of having bivariate data. However, it seems that having temperature data alone does a substantially worse job in terms of explained variance

>>>>

Note that this strategy gives a sequential conditional version of
expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
the expectation is conditional to this data. A similar closed-form calculation
for EIBV is hence applicable in Eq.
\eqref{critSEQ} using the
updated GP model from stage $s-1$. Once the data is collected for the
best design at stage $s$, the GP model is updated again. The mean $\bbm^{s}$ and
covariance matrix $\Sigma^{s}$ are used to compute the next design at
stage $s+1$, and so on.

@@

We now extend the myopic strategy to a look-ahead strategy, which is
optimal when one can gather only two more stages of measurements. In
addition to the next stage of measurements $\by^{j,s}$, this
look-ahead strategy, anticipates the subsequent design indexed $j_2$ with
data $\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$.  The
selected design is:
\begin{eqnarray}\label{critLA}
    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( \mbox{argmin}_{j_2}  V_{j_2} \right) \right\}, \\
V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1}) ( 1-p(\x;\mathcal{Y}^{j,s+1}) ) \right) d\mes(u), \nonumber \\
    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
\end{eqnarray}
Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$ represent
the sets of data variables, and the expectations are calculated with respect to
the conditional densities given data at that stage, meaning we have two levels of intermixed minimizations and integrals.
We solve the
first expectation by Monte Carlo sampling of data $\by_{d^{j,s}}$ from its
conditional distribution. For each of these data samples, the second expectation
is solved using the closed-form expressions for EIBV from
Section \ref{sec:ESEP}.

@@

For a fixed survey length, a closed-form expression for the EIBV is available as in Eq. \eqref{eq:eibv}. However, for the sequential approaches this is not the
case. For comparison the properties are therefore evaluated using
Monte Carlo sampling over several replicates of realizations from the model while conducting simulated sequential surveys for each
one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
gathered along the north-south survey lines in these displays. In the
Monte Carlo replicates, such results are repeatedly computed to
approximate the EIBV. We also compare predictive performance
measured by root mean square error (RMSE) for temperature and salinity
estimates as well as also the variance reduction in these two variables. It is
important to note that the objective function used by the agent is
focused on reducing the EIBV, but we nevertheless expect that
we will achieve good predictive performance for criteria such as RMSE
as well. Another non-statistical criteria that is important for
practical purposes is the computational time needed for the strategy,
as this will impact the performance for an embedded system.

@@

We studied the sensitivity of the results by modifying the input parameters to have different correlations between temperature and salinity, standard deviations, and spatial correlation range.  In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of realized IBV, and much better than \textit{naive}. The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small initial standard deviation or very large spatial correlation range. The \textit{static\_north} design continues to be the best static design for IBV, while \textit{static\_zigzag} is the best design for the other predictive performance measures, especially so with large spatial correlation range. We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the IBV results are not much worse when only temperature data are available. In addition to the comparison made in Table \ref{tab:sim_rhoab}, the current setting includes spatial correlation and this likely reduces the additional influence of having bivariate data. However, it seems that having temperature data alone does a substantially worse job in terms of explained variance
